---
title: "P8105 - Homework 1"
author: "Mingkuan Xu"
output: html_document
---

### Problem 1

Create a data frame comprised of:

- a random sample of size 10 from a standard Normal distribution
- a logical vector indicating whether elements of the sample are greater than 0
- a character vector of length 10
- a factor vector of length 10, with 3 different factor “levels”

```{r p1_create}

set.seed(124) # Set a random seed

# Generate a random normal distributed sample of size 10
sample <- rnorm(10)

# Add the sample to a dataframe with column name random_values
df <- data.frame(random_values = sample) 

# Add a new column named is_positive indicating whether elements of the sample are greater than 0
df$is_positive <- with(df,ifelse(random_values>0,TRUE,FALSE))

# Create a character vector of length 10 and name it character_vector
df$character_vector = vector("character", length = 10)

# Create a factor vector with 3 different factor “levels”, and name it factor_vector
df$factor_vector = factor(c(1,2,2,3,1,2,3,3,1,2))

```

Try to take the mean of each variable in your dataframe. What works and what doesn’t?

Hint: to take the mean of a variable in a dataframe, you need to pull the variable out of the dataframe. Try loading the tidyverse and using the pull function.

```{r p1_find_means}

# We can use the pull function from tidyverse as instructed.

library(tidyverse) # Load tidyverse library

mean(pull(df,random_values)) # Find the mean of numeric vector
mean(pull(df,is_positive)) # Find the mean of logical vector
mean(pull(df,character_vector)) # Find the mean of character vector
mean(pull(df,factor_vector)) # # Find the mean of factor vector
```

**By running the code, we could see it works for the numeric and logical vectors, but fails for the character and factor vectors.**

In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). 

```{r p1_as_numeric,results = 'hide'}
as.numeric(df$is_positive)
as.numeric(df$character_vector)
as.numeric(df$factor_vector)
```
What happens, and why? Does this help explain what happens when you try to take the mean?

**For the logical vector, as.numeric transforms it to a vector with values 0 (for FALSE) or 1 (for TRUE), which explains how R is able to find the mean of this vector.**

**For the character vector, as.numeric transforms it to multiple NAs. This also explains why there cannot be a "mean" value of character vecter, because characters cannot be automatically transformed into numbers in R.** 

**For the factor vector, as.numeric does transform it to a numeric vector. However, we still cannot directly apply the mean functiion on a factor vector. From the warning message we seen above, "argument is not numeric or logical: returning NA", it looks like the mean value function only takes numeric or logical inputs.**


### Problem 2

Use the following code (in your RMarkdown file) to load the penguins dataset:

```{r p2_load_data}
data("penguins", package = "palmerpenguins")
```

Write a short description of the penguins dataset (not the penguins_raw dataset) using inline R code, including:
- The data in this dataset, including names / values of important variables
- The size of the dataset (using nrow and ncol)
- The mean flipper length


**This dataset describes the information of some penguins, including their `r names(penguins)`. Typical species includes `r unique(penguins$species)`. Typical islands found includes `r unique(penguins$island)`. The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns. After removing all the NA valules, the mean flipper length is `r mean(penguins$flipper_length_mm,na.rm=TRUE)` mm.**

Make a scatterplot of flipper_length_mm (y) vs bill_length_mm (x); color points using the species variable (adding color = ... inside of aes in your ggplot code should help).

```{r p2_plot}
ggplot(penguins,aes(x=bill_length_mm, y=flipper_length_mm,color=species))+
    geom_point()+
  labs(x="Bill Length (mm)",y="Flipper Length (mm)")
```
Export your first scatterplot to your project directory using ggsave.

```{r p2_save_plot}
ggsave(file="./output.pdf")
```
